---
title: "zjin26_mini-assign3"
author: "Vincent Jin"
date: "2023-04-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ============================================================ 
# 	Mini-assignment # 3
# ============================================================ 

## Vincent Jin

# (1) create a vector containing numbers 0 to 10, 100 to 200, and 900 to 1000 (without entering all numbers individually)

``` {r}
v <- c(0:10, 100:200, 900:1000)
```

# (2) find the min, max, sum, average and quartiles of the last vector

``` {r}
min(v)
max(v)
sum(v)
mean(v)
quantile(v)
```

# (3) find the min, max, sum, average and quartiles of the following vector: c(rnorm(1000,500,10))  [note that rnorm command is generating random numbers every time you run it!]

``` {r}
v1 <- c(rnorm(1000,500,10))
min(v1)
max(v1)
sum(v1)
mean(v1)
quantile(v1)
```

# (4) create a data.frame with four columns (age, sex, cost, risk) with 1000 rows, as follow: 
# first column randomly picked from 18 to 65; 
# second column random 'F' or 'M' or NA (i.e., missing); 
# third column with random numbers between 1000 and 100,000 (round to 1 digit); and, 
# fourth column with random number between 0 and 1 (do NOT round)

``` {r}
df1 <- data.frame(age = sample(18:65, 1000, replace = T),
           sex = sample(c('M','F', NA), 1000, replace = T),
           cost = round(runif(1000, min = 1000, max = 1000000), 1),
           risk = runif(1000, min = 0, max = 1))
```

# (5) assign the previous data.frame to a variable such as x
# get the mean of all columns in the data.frame
# count the number of F and M and NA (find online how to count NA in the table command)
# set your seed number to 100

``` {r}
set.seed(100)
x <- data.frame(age = sample(18:65, 1000, replace = T),
           sex = sample(c('M','F', NA), 1000, replace = T),
           cost = round(runif(1000, min = 1000, max = 1000000), 1),
           risk = runif(1000, min = 0, max = 1))
col <- names(x)
for (i in col) {
  if (typeof(x[[i]]) != "character") {  
    cat(paste("the mean of", i, "is:\n"))
    print(mean(x[[i]]))
  }
}
table(x$sex, useNA = "always")
```

# (6) get the quartiles of the cost and risk columns before and after log transformation

``` {r}
col <- c('cost', 'risk')
for (i in col) {
  cat(paste("the quartiles of", i, "is:\n"))
  print(quantile(x[[i]]))
  cat(paste("the quartiles of", i, "after log transformation is:\n"))
  print(quantile(log(x[[i]])))
}
```

# (7) count how many of the costs are more than mean of cost 
# avoid entering a plain number in the comparison

``` {r}
sum(x$cost > mean(x$cost))
```